---
title: "motus_viewshed"
author: "fork of code by adamdsmith on GitHub"
date: "10/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Source

https://github.com/adamdsmith/motus_viewshed

I forked this repo and then cloned it into /d/PatrickLorch/MotusGeneral/rmotus_work with 
`git clone https://github.com/southern-sierra-research/motus_viewshed.git`

```{r cars}
source("D:/PatrickLorch/MotusGeneral/rmotus_work/motus_viewshed/R/motus_viewshed.R")
source("D:/PatrickLorch/MotusGeneral/rmotus_work/motus_viewshed/R/motus_prop_viewshed_visible.R")

# Example for Red Tail Hill in Kern River valley
# RTH motus_prop_viewshed_visible(coords= c(35.6837, -118.3053), ht = 10)
motus_prop_viewshed_visible(coords= c(37.176130, -120.802421), ht = 10) # San Luis NWR
motus_viewshed(coords= c(37.176130, -120.802421), ht = 10) # San Luis NWR
motus_prop_viewshed_visible(coords= c(35.478637, -118.211096), ht = 10) # Tunnel Spring
motus_viewshed(coords= c(35.478637, -118.211096), ht = 10) # Tunnel Spring
motus_prop_viewshed_visible(coords= c(37.176130, -120.802421), ht = 10) # San Luis NWR
motus_viewshed(coords= c(37.176130, -120.802421), ht = 10) # San Luis NWR
motus_prop_viewshed_visible(coords= c(37.176130, -120.802421), ht = 10) # San Luis NWR
motus_viewshed(coords= c(37.176130, -120.802421), ht = 10) # San Luis NWR
motus_prop_viewshed_visible(coords= c(37.176130, -120.802421), ht = 10) # San Luis NWR
motus_viewshed(coords= c(37.176130, -120.802421), ht = 10) # San Luis NWR

I still need to limit this to just the area that is not obscured at that point, 
but doing viewsheds in R is tougher than I thought.

```


## Viewshed package testing

What I tried but did not use.

### Rayshader

https://www.rayshader.com/

Does not work because it fixes angle and compas direction of sun.

```{r rayshader}
# To install the latest version from Github:
# install.packages("devtools")
devtools::install_github("tylermorganwall/rayshader")
library(rayshader)
library(dplyr)
```


### r.viewshed in GRASS

This will work, but we have to figure out how to launch GRASS from within R

https://grasswiki.osgeo.org/wiki/R_statistics/rgrass7

Does not work unless you already have a grass workspace and database set up.

```{r grass}
# install.packages("rgrass7")
library(rgrass7)

initGRASS(gisBase = "C:\Program Files\QGIS 3.20.3\bin\grass78.bat",
         gisDbase = "C:\Users\PatrickLorch\Documents\grassdata",
         location = "nc_spm_08_grass7", 
         mapset = "user1", 
         SG="elevation")

```

### stackoverflow solution

https://stackoverflow.com/questions/21841387/r-code-that-evaluates-line-of-sight-los-between-two-lat-lon-points

This requires the xy coordinates to be in easting and northing for pythagorean
theorum to be used.

```{r viewshed}
# Get a raster of elevations
coords = c(35.6837, -118.3053)
ht = 10
zoom = 10
lon <- coords[2]; lat <- coords[1]
coords <- cbind(lon, lat)
prj <- sprintf("+proj=laea +x_0=0 +y_0=0 +lon_0=%f +lat_0=%f", lon, lat)

pt_ll <- sp::SpatialPoints(coords, sp::CRS("+init=epsg:4326"))
pt <- sp::spTransform(pt_ll, sp::CRS(prj))
poly20 <- rgeos::gBuffer(pt, width = 20000, quadsegs = 12)

# Get DEM
suppressWarnings(
  suppressMessages(
    elev <- elevatr::get_elev_raster(poly20, z = zoom, clip = "locations", verbose = FALSE)
  )
)

rasterprofile <- function(r, xy1, xy2){
### sample a raster along a straight line between two points
### try to match the sampling size to the raster resolution
    dx = sqrt( (xy1[1]-xy2[1])^2 + (xy1[2]-xy2[2])^2 )
    nsteps = 1 + round(dx/ min(res(r)))
    xc = xy1[1] + (0:nsteps) * (xy2[1]-xy1[1])/nsteps
    yc = xy1[2] + (0:nsteps) * (xy2[2]-xy1[2])/nsteps
    data.frame(x=xc, y=yc, z=r[cellFromXY(r,cbind(xc,yc))])
}

cansee <- function(r, xy1, xy2, h1=0, h2=0){
### can xy1 see xy2 on DEM r?
### r is a DEM in same x,y, z units
### xy1 and xy2 are 2-length vectors of x,y coords
### h1 and h2 are extra height offsets
###  (eg top of mast, observer on a ladder etc)
    xyz = rasterprofile(r, xy1, xy2)
    np = nrow(xyz)-1
    h1 = xyz$z[1] + h1
    h2 = xyz$z[np] + h2
    hpath = h1 + (0:np)*(h2-h1)/np
    return(!any(hpath < xyz$z))
}

viewTo <- function(r, xy, xy2, h1=0, h2=0, progress="none"){
    ## xy2 is a matrix of x,y coords (not a data frame)
    require(plyr)
    aaply(xy2, 1, function(d){cansee(r,xy,d,h1,h2)}, .progress=progress)
}



```

## View-cone proportion

If the viewshed is represented as a cone, with the antenna direction being the 
center of the cone, a verticle cross section is an isosceles triangle and an 
approximation of the viewshed in any given direction.  We calculate the width 
of the cone at each distance (h) as an approximation of the potential viewshed 
at any distance (2*b below).  If the difference between the antenna elevation 
and the pixel elevation (de) is less than 2*b, we can use (2*b - de)/ 2*b to 
represent the fraction of the potential viewshed at each point.  

To improve this estimate, we can represent the potential viewshed as something
more like the actual antenna sensetivity.  So we could replace the current
linear function of distance from the antenna, we can use one that initially
increased with distance and then decreases with further distance.  This same
function can be used to weight the estimate to account for the 3D shape of the
detection bubble.

The other enhancement that would be nice would be to use rayshader package to 
do the above calculations only at points not blocked from view.

Yet another enhancement would be to find other points that would give a better
fraction of the the potential viewshed in a given direction or set of
directions.

For now we use `halfangle` to represent how wide the cone is (2*halfangle).

```{r vcprop}
vcprop = function(halfangle = 10, antht, h){
# halfangle = half of the angle of view of antenna detection cone
# h = distance from antenna
# pe = pixel elevation, 
# antht = antenna height above sea level (elevation + height above ground)
# all in meters
  b = tan(halfangle) * h
  de = pe - antht
  if(de < 2*b)
    fpvs = (2*b - de)/ 2*b
  else
    fpvs = 0
  return(fpvs)
}

```

